---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load the data then convert the date column to date format.
```{r}
activity <- read.csv('activity.csv')
activity$date <- as.Date(activity$date)
```


## What is mean total number of steps taken per day?
```{r}
step_date <- aggregate(steps~date, data = activity, sum)
library(ggplot2)
ggplot(data = step_date, aes(x=steps)) + geom_histogram()
mean_step <- mean(step_date$steps)
median_step <- median(step_date$steps)
```
The mean and median of the total number of steps taken per day are `r sprintf("%.2f", mean_step)` and `r sprintf("%.2f", median_step)` respectively.

## What is the average daily activity pattern?
```{r}
step_interval <- aggregate(steps~interval, data = activity, sum)
ggplot(step_interval, aes(x=interval, y=steps)) + geom_line()

max_step <- step_interval[which.max(step_interval$steps),]$interval
```
5-minute interval `r max_step`, on average across all the days in the dataset, contains the maximum number of steps

## Imputing missing values
```{r}
total_missing <- sum(is.na(activity$steps))
```
There are `r total_missing` missing values in the data.
```{r}
library(dplyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))

activity_impute <- activity %>%
  group_by(interval) %>%
  mutate(
    steps = impute.mean(steps)  
  )
step_date_impute <- aggregate(steps~date, data = activity_impute, sum)
ggplot(data = step_date_impute, aes(x=steps)) + geom_histogram()
mean_impute <- mean(step_date_impute$steps)
median_impute <- median(step_date_impute$steps)
```
The missing values are imputed by the mean for that 5-minute interval. The histogram with imputed values has higher peak. 

The mean and median of the total number of steps taken per day are `r sprintf("%.2f", mean_impute)` and `r sprintf("%.2f", median_impute)` respectively.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity$weekend <- 'weekend'
activity[weekdays(activity$date) %in% c('Saturday', 'Sunday'),][['weekend']] <- 'weekday'
names(activity)
ggplot(data = activity, aes(interval, steps)) + geom_line() + facet_grid(weekend~.)
```

At the weekend, the total of steps is higher.